---
title: "Intermediate presentation"
subtitle: "Differential regulation"
author: "Joël Meili"
date: "7/6/2021"
output: 
  beamer_presentation:
    theme: "Boadilla"
    colortheme: "whale"
    fonttheme: "structurebold"
    slide_level: 2
bibliography: literature/intermediate_presentation.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

suppressPackageStartupMessages({
  library(png)
  library(rasterpdf)
  library(grid)
  library(gridExtra)
})
```

## Contents

- Brief (re)-introduction to the topic
- Data selection process
- Explorative data analysis of spliced and unspliced counts
- Outlook

## (Re)-Introduction to the topic

- Abundance of spliced and unspliced mRNA methods allows studying cellular dynamical processes, such as cell differentiation, cell cycle and cell regulation
- Aim at discovering differential usage of spliced and unspliced counts between groups of samples in each cluster of cells

## Data selection process

- Mouse data:
  - kidney cells from four biological replicates under the same experimental condition (@mouse)
  
- Human data:
  - 16 brain organoids (PGP1 cell line) from three different batches at three months and six months growth (different experimental condition) (@human)

# Mouse data

## Looking at the proportion of spliced and unspliced counts

```{r out.width = "100%"}
path <- "../RNAVelo/kidney_mouse/04_figures/"

normal1 <- readPNG(source = paste0(path, "individual_samples/normal1/proportions.png"))
normal2 <- readPNG(source = paste0(path, "individual_samples/normal2/proportions.png"))
normal3 <- readPNG(source = paste0(path, "individual_samples/normal3/proportions.png"))
normal4 <- readPNG(source = paste0(path, "individual_samples/normal4/proportions.png"))

grid.arrange(
  arrangeGrob(rasterGrob(normal1, height = 0.7,
                         width = 0.95), top = "Normal 1"),
  arrangeGrob(rasterGrob(normal2, height = 0.7,
                         width = 0.95), top = "Normal 2"),
  arrangeGrob(rasterGrob(normal3, height = 0.7,
                         width = 0.95), top = "Normal 3"), 
  arrangeGrob(rasterGrob(normal4, height = 0.7,
                         width = 0.95), top = "Normal 4"), ncol = 2)
```

## Looking at the RNA velocities - individual samples

```{r out.width = "100%"}
normal1 <- readPNG(source = paste0(path, "individual_samples/normal1/dynamical_model.png"))
normal2 <- readPNG(source = paste0(path, "individual_samples/normal2/dynamical_model.png"))
normal3 <- readPNG(source = paste0(path, "individual_samples/normal3/dynamical_model.png"))
normal4 <- readPNG(source = paste0(path, "individual_samples/normal4/dynamical_model.png"))

grid.arrange(
  arrangeGrob(rasterGrob(normal1), top = "Normal 1"),
  arrangeGrob(rasterGrob(normal2), top = "Normal 2"),
  arrangeGrob(rasterGrob(normal3), top = "Normal 3"), 
  arrangeGrob(rasterGrob(normal4), top = "Normal 4"), ncol = 2)
```


## Looking at the RNA velocities - all samples

```{r out.width = "100%"}
normal1 <- readPNG(source = paste0(path, "all_samples/normal1/dynamical_model.png"))
normal2 <- readPNG(source = paste0(path, "all_samples/normal2/dynamical_model.png"))
normal3 <- readPNG(source = paste0(path, "all_samples/normal3/dynamical_model.png"))
normal4 <- readPNG(source = paste0(path, "all_samples/normal4/dynamical_model.png"))

grid.arrange(
  arrangeGrob(rasterGrob(normal1), top = "Normal 1"),
  arrangeGrob(rasterGrob(normal2), top = "Normal 2"),
  arrangeGrob(rasterGrob(normal3), top = "Normal 3"), 
  arrangeGrob(rasterGrob(normal4), top = "Normal 4"), ncol = 2)
```

# Human data
## Looking at the proportion of spliced and unspliced counts

```{r out.width = "100%"}
path <- "../RNAVelo/brain_human/04_figures/"

organoid1 <- readPNG(source = paste0(path, "individual_samples/organoid1/proportions.png"))
organoid2 <- readPNG(source = paste0(path, "individual_samples/organoid2/proportions.png"))
organoid3 <- readPNG(source = paste0(path, "individual_samples/organoid3/proportions.png"))
organoid16 <- readPNG(source = paste0(path, "individual_samples/organoid16/proportions.png"))
organoid17 <- readPNG(source = paste0(path, "individual_samples/organoid17/proportions.png"))
organoid18 <- readPNG(source = paste0(path, "individual_samples/organoid18/proportions.png"))

grid.arrange(
  arrangeGrob(rasterGrob(organoid1, height = 0.7,
                         width = 0.95), top = "Organoid 1"),
  arrangeGrob(rasterGrob(organoid2, height = 0.7,
                         width = 0.95), top = "Organoid 2"),
  arrangeGrob(rasterGrob(organoid3, height = 0.7,
                         width = 0.95), top = "Organoid 3"), 
  arrangeGrob(rasterGrob(organoid16, height = 0.7,
                         width = 0.95), top = "Organoid 16"),
  arrangeGrob(rasterGrob(organoid17, height = 0.7,
                         width = 0.95), top = "Organoid 17"),
  arrangeGrob(rasterGrob(organoid18, height = 0.7,
                         width = 0.95), top = "Organoid 18"), ncol = 3)
```

## Looking at the RNA velocities - individual samples

```{r out.width = "100%"}
organoid1 <- readPNG(source = paste0(path, "individual_samples/organoid1/dynamical_model.png"))
organoid2 <- readPNG(source = paste0(path, "individual_samples/organoid2/dynamical_model.png"))
organoid3 <- readPNG(source = paste0(path, "individual_samples/organoid3/dynamical_model.png"))
organoid16 <- readPNG(source = paste0(path, "individual_samples/organoid16/dynamical_model.png"))
organoid17 <- readPNG(source = paste0(path, "individual_samples/organoid17/dynamical_model.png"))
organoid18 <- readPNG(source = paste0(path, "individual_samples/organoid18/dynamical_model.png"))

grid.arrange(
  arrangeGrob(rasterGrob(organoid1), top = "Organoid 1"),
  arrangeGrob(rasterGrob(organoid2), top = "Organoid 2"),
  arrangeGrob(rasterGrob(organoid3), top = "Organoid 3"), 
  arrangeGrob(rasterGrob(organoid16), top = "Organoid 16"),
  arrangeGrob(rasterGrob(organoid17), top = "Organoid 17"),
  arrangeGrob(rasterGrob(organoid18), top = "Organoid 18"), ncol = 3)
```

## Looking at the RNA velocities - all samples

```{r out.width = "100%"}
organoid1 <- readPNG(source = paste0(path, "all_samples/organoid1/dynamical_model.png"))
organoid2 <- readPNG(source = paste0(path, "all_samples/organoid2/dynamical_model.png"))
organoid3 <- readPNG(source = paste0(path, "all_samples/organoid3/dynamical_model.png"))
organoid16 <- readPNG(source = paste0(path, "all_samples/organoid16/dynamical_model.png"))
organoid17 <- readPNG(source = paste0(path, "all_samples/organoid17/dynamical_model.png"))
organoid18 <- readPNG(source = paste0(path, "all_samples/organoid18/dynamical_model.png"))

grid.arrange(
  arrangeGrob(rasterGrob(organoid1), top = "Organoid 1"),
  arrangeGrob(rasterGrob(organoid2), top = "Organoid 2"),
  arrangeGrob(rasterGrob(organoid3), top = "Organoid 3"), 
  arrangeGrob(rasterGrob(organoid16), top = "Organoid 16"),
  arrangeGrob(rasterGrob(organoid17), top = "Organoid 17"),
  arrangeGrob(rasterGrob(organoid18), top = "Organoid 18"), ncol = 3)
```

## Conclusions from the explorative analysis

- We expected comparable patterns with changing trajectories between experimental conditions
- No clear patterns that were consistent between biological replicates, also arises the question of reproducibility
- Interpretation based on RNA velocity seems not possible

## Simulation of data sets from the mouse data

1. Compute pseudobulk from the four mouse samples and filter for these cell types:
**Epithelial cells, Adipocytes and Hepatocytes**

2. For each of the cell types randomly draw a fixed proportion of genes and cells. Exchange the counts of spliced and unspliced counts for the drawn genes and cells in one experimental condition.

4. Differential analysis using eisaR and DEXSeq -> Establishing baseline performance

5. Adding multimapping with minnow -> We expect performance of eisaR and DEXSeq to go down, which should improve again with our proposed method

## References


